---
title: "Normal and Other Continuous distributions"
author: "Manideep Bangaru"
date: "28/11/2020"
output: html_document
---
\

Continuous probability distributions vary by the shape under the curve.

### Normal Distribution
The normal distribution is not only symmetrical, but bell-shaped, a shape that (loosely) suggests the profile of a bell. Being bell-shaped means that most values of the continuous vari- able will cluster around the mean.

### Uniform Distribution
It is also known as rectangular distribution, contains the values that are equally distributed in the range between the minimum value and the maximum value.
In a uniform distribution every value has a equal chance of occurrence

### Exponential Distribution
The exponential distribution contains values from zero to positive infinity and is right-skewed, making the mean greater than the median

### Properties of Normal Distribution
* Symmetrical distribution. Its mean and median are therefore equal
* Bell-shaped. Values cluster around the mean
* Interquartile range is roughly 1.33 standard deviations. Therefore, the middle 50% of the values are contained within an interval that is approximately two-thirds of a standard deviation below and two-thirds of a standard deviation above the mean
* The distribution has an infinite range $-\infty < X < \infty$
* Six standard deviations approximate this range

### Normal Probability Density Function
$\large f(x) = \frac{1}{\sigma\sqrt{2\pi}}e^-(\frac{1}{2})(\frac{x-\mu}{\sigma})^2$ \
\
where,\
e = mathematical constant (2.71828) \
$\pi$ = mathematical constant (3.14159) \
$\mu$ = mean \
$\sigma$ = standard deviation \
X = Any value of the continuous variable, $-\infty < X < \infty$ \

### Computing Normal Probabilities
To compute Normal probabilites, you need to first convert **X** to a standardized normal variable **(Z)**, using <font color='red'> Transformation formula </font>

### Z Transformation formula
The Z value is equal to difference between X and the mean,$\mu$, divided by the standard deviation($\sigma$)

$\text{Z =} \large \frac{x-\mu}{\sigma}$

### Excercises
1) What is the probability that the load time of a web sales page will be more than 9 seconds if average load time is 7 seconds with a standard deviation of 2?
```{r, echo=TRUE}
1-pnorm(9,7,2)
```

2) What is the probability that the load time for the web sales page will be less than 7 seconds or more than 9 seconds?
```{r, echo=TRUE}
pnorm(7,7,2) + (1-pnorm(9,7,2))
```

3) How much time (in seconds) will elapse before the fastest 10% of the web sales pages load time occur?
```{r, echo=TRUE}
qnorm(0.10,7,2)
```
_Thus, 10% of the load times are 4.44 seconds or less_

### Evaluating Normality
* Construct charts and observe their appearance. For small or moderate sized data sets, create a stem-and-leaf display or a boxplot. For large data sets, in addition, plot a histogram or polygon.

* Compute descriptive statistics and compare these statistics with the theoretical proper- ties of the normal distribution. Compare the mean and median. Is the interquartile range approximately 1.33 times the standard deviation? Is the range approximately 6 times the standard deviation?

* Evaluate how the values are distributed. Determine whether approximately two-thirds of the values lie between the mean and $\pm$ 1 standard deviation. Determine whether approximately four-fifths of the values lie between the mean and $\pm$ 1.28 standard deviations. Determine whether approximately 19 out of every 20 values lie between the mean and $\pm$ 2 standard deviations.

### The Uniform Distribution
In the uniform distribution, the values are evenly distributed in the range between the smallest value, **a**, and the largest value, **b**. 

Because of its shape, the uniform distribution is sometimes called the **rectangular distribution**

**_Uniform Probability Density Function_**

$\large f(x) = \Large \frac{1}{b-a}, \normalsize \text{if a} \le \text{X} \le b \land \text{0 elsewhere}$

where,\
a = minimum value of X \
b = maximum value of X