---
title: "Numerical Descriptive Measures"
author: "Manideep Bangaru"
date: "21/11/2020"
output: html_document
---
\
$\Large \sf To \ clean \ the \ environment \ variables$

```{r, echo=TRUE}
rm(list=ls(all=T))
```
\
$\Large \sf Defining \ a \ Vector$
```{r,echo=T}
vec <- c(20,34,56,23,45,22,60,23,56,78,23,45)
```
\
$Llarge \sf Mean \ of \ the \ vector$\

$\frac{\Large\sum_{\Large i=1}^{\Large n}{\Large x_i}}{\Large n}$
```{r,echo=TRUE}
mean(vec)
```
\
$\Large \sf Median\ of \ the \ vector$\

$\Large \frac{(n+1)}{2} \normalsize \sf \  ranked \ value$
```{r,echo=TRUE}
median(vec)
```

* For symmetrical distributed data mean, median and mode are almost equal in value
* For asymmetrical distributed data, following relationship holds good approximately
* Mode = 3 * Median - 2 * mean (or)
* Mean - Mode = 3 * (Mean - Median)
* above realtion is called as empirical relation. Using this if two measures are known, it is easy to find out the third measure

\
$\Large \sf Harmonic\ Mean\ of\ a\ vector$

$\Large\frac {\Large n}{\Large \sum_{\Large i=1}^{\Large n}\Large \frac{\Large 1}{\Large x_i}}$
```{r, echo=TRUE}
hm_ <- function(vcetor) {
  return(length(vec)/sum(1/vec))
}
```
```{r,echo=TRUE}
hm_(vec)
```
\
$\Large \sf Geometric\ Mean\ of\ a\ vector$

$\large \bar X_g = (X_1*X_2*X_3 * \ .... \ X_n )^\frac{1}{n}$
```{r,echo=T}
gm_ <- function(vcetor) {
  return(prod(vcetor)^(1/length(vcetor)))
}
```
```{r,echo=T}
gm_(vec)
```

\
\
$\sf \huge Variation\ and\ shape$
\
\
$\sf \Large Range\ of\ a\ vector$\
\
$\sf \normalsize Range = (\large x_{max}-\large x_{min})$
```{r,echo=T}
range_ <- function(vcetor) {
  return(max(vcetor)-min(vcetor))
}
```
```{r,echo=T}
range_(vec)
```
\
$\sf \Large Variance\ of\ a\ vector$\
$\Large \sigma^2 \small =\Large \frac{\sum_{i=1}^{n} (x_i-\bar X)^2}{n}$
```{r,echo=T}
variance_ <- function(vcetor){
    var_ <-  0
    for (ele in 1:length(vcetor)){
      var_ = var_ + (vcetor[ele] - mean(vcetor))^2
    }
    return (var_/length(vcetor)-1)
}
```
```{r,echo=T}
variance_(vec)
```
\
$\sf \Large Standard\ deviation\ of\ a\ vector$

$\Large \sigma \small =\sqrt{\Large \frac{\sum_{i=1}^{n} (x_i-\bar X)^2}{n}}$
```{r,echo=T}
stddev_ <- function(vcetor){
  return (variance_(vcetor) ^ 0.5)
}
```
```{r,echo=T}
stddev_(vec)
```
\
$\sf \Large Coefficient\ of\ Variation$
\
It measures the scatter in the data with respect to the mean
\
\
$\Large (\frac {\sigma}{\mu})*100$
```{r,echo=T}
CoeffVar_ <- function(vcetor){
  return (stddev_(vcetor)/mean(vcetor)) *100
}
```
```{r,echo=T}
CoeffVar_(vec)
```
\
$\sf \Large Z-Score$
\
\
$\Large(\frac{x-\mu}{\sigma})$
\

* a Z score of 0 indicates that the value is same as the mean
* Helps in identifying outliers, less than -3 and greater than +3 are considered to be outliers
```{r,echo=T}
zscore <- function(vcetor){
  temp_ <- c()
  for (ele in 1:length(vcetor)){
    temp_ <- append(temp_, (vcetor[ele]-mean(vcetor))/stddev_(vcetor))
  }
  return (temp_)
}
```
```{r,echo=T}
zscore(vec)
```
\
$\Large Skewness$
\

* Mean < Median ----> left skewed or negative skew
* Mean = Median ----> symmetrical distribution (zero skewness)
* Mean > Median ----> right skewed or positive skew

\
$\Large Kurtosis$
\

* A distribution that has a sharper-rising center peak than the peak of a normal distribution has positive kurtosis, a kurtosis value that is greater than zero, and is called **lepokurtic** 
* A distribution that has a slower-rising (flatter) center peak than the peak of a normal distribution has negative kurtosis, a kurtosis value that is less than zero, and is called **platykurtic**
* A **lepokurtic distribution** has a **higher concentration** of values near the mean of the distribution compared to a normal distribution, while a **platykurtic distribution** has a **lower concentration** compared to a normal distribution

